### 2021-05-07 FMCAD2021/paper.tex
change to IEEE Trans template

% !TEX program = xelatex
%
\documentclass[10pt,letterpaper]{IEEEtran}

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % The IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%\usepackage{acronym}
% acronym.sty was written by Tobias Oetiker. This package provides tools for
% managing documents with large numbers of acronyms. (You don't *have* to
% use this package - unless you have a lot of acronyms, you may feel that
% such package management of them is bit of an overkill.)
% Do note that the acronym environment (which lists acronyms) will have a
% problem when used under IEEEtran.cls because acronym.sty relies on the
% description list environment - which IEEEtran.cls has customized for
% producing IEEE style lists. A workaround is to declared the longest
% label width via the IEEEtran.cls \IEEEiedlistdecl global control:
%
% \renewcommand{\IEEEiedlistdecl}{\IEEEsetlabelwidth{SONET}}
% \begin{acronym}
%
% \end{acronym}
% \renewcommand{\IEEEiedlistdecl}{\relax}% remember to reset \IEEEiedlistdecl
%
% instead of using the acronym environment's optional argument.
% The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/acronym

%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/pkg/mdwtools

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/pkg/eqparbox

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/pkg/endfloat
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% NOTE: PDF thumbnail features are not required in IEEE papers
%       and their use requires extra complexity and work.
%\ifCLASSINFOpdf
%  \usepackage[pdftex]{thumbpdf}
%\else
%  \usepackage[dvips]{thumbpdf}
%\fi
% thumbpdf.sty and its companion Perl utility were written by Heiko Oberdiek.
% It allows the user a way to produce PDF documents that contain fancy
% thumbnail images of each of the pages (which tools like acrobat reader can
% utilize). This is possible even when using dvi->ps->pdf workflow if the
% correct thumbpdf driver options are used. thumbpdf.sty incorporates the
% file containing the PDF thumbnail information (filename.tpm is used with
% dvips, filename.tpt is used with pdftex, where filename is the base name of
% your tex document) into the final ps or pdf output document. An external
% utility, the thumbpdf *Perl script* is needed to make these .tpm or .tpt
% thumbnail files from a .ps or .pdf version of the document (which obviously
% does not yet contain pdf thumbnails). Thus, one does a:
% 
% thumbpdf filename.pdf 
%
% to make a filename.tpt, and:
%
% thumbpdf --mode dvips filename.ps
%
% to make a filename.tpm which will then be loaded into the document by
% thumbpdf.sty the NEXT time the document is compiled (by pdflatex or
% latex->dvips->ps2pdf). Users must be careful to regenerate the .tpt and/or
% .tpm files if the main document changes and then to recompile the
% document to incorporate the revised thumbnails to ensure that thumbnails
% match the actual pages. It is easy to forget to do this!
% 
% Unix systems come with a Perl interpreter. However, MS Windows users
% will usually have to install a Perl interpreter so that the thumbpdf
% script can be run. The Ghostscript PS/PDF interpreter is also required.
% See the thumbpdf docs for details. The latest version and documentation
% can be obtained at.
% http://www.ctan.org/pkg/thumbpdf

% NOTE: PDF hyperlink and bookmark features are not required in IEEE
%       papers and their use requires extra complexity and work.
% *** IF USING HYPERREF BE SURE AND CHANGE THE EXAMPLE PDF ***
% *** TITLE/SUBJECT/AUTHOR/KEYWORDS INFO BELOW!!           ***
\newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
pdftitle={A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints},%<!CHANGE!
pdfsubject={String constraint solving},%<!CHANGE!
pdfauthor={Hao Wu, Yu-Fang Chen, Zhilin Wu, Naijun Zhan},%<!CHANGE!
pdfkeywords={String-integer conversion, Flat regular constraints, Presburger arithmetic, Exponential function, Quantifier elimination.}}%<^!CHANGE!
%\ifCLASSINFOpdf
%\usepackage[\MYhyperrefoptions,pdftex]{hyperref}
%\else
%\usepackage[\MYhyperrefoptions,breaklinks=true,dvips]{hyperref}
%\usepackage{breakurl}
%\fi
% One significant drawback of using hyperref under DVI output is that the
% LaTeX compiler cannot break URLs across lines or pages as can be done
% under pdfLaTeX's PDF output via the hyperref pdftex driver. This is
% probably the single most important capability distinction between the
% DVI and PDF output. Perhaps surprisingly, all the other PDF features
% (PDF bookmarks, thumbnails, etc.) can be preserved in
% .tex->.dvi->.ps->.pdf workflow if the respective packages/scripts are
% loaded/invoked with the correct driver options (dvips, etc.). 
% As most IEEE papers use URLs sparingly (mainly in the references), this
% may not be as big an issue as with other publications.
%
% That said, Vilar Camara Neto created his breakurl.sty package which
% permits hyperref to easily break URLs even in dvi mode.
% Note that breakurl, unlike most other packages, must be loaded
% AFTER hyperref. The latest version of breakurl and its documentation can
% be obtained at:
% http://www.ctan.org/pkg/breakurl
% breakurl.sty is not for use under pdflatex pdf mode.
%
% The advanced features offer by hyperref.sty are not required for IEEE
% submission, so users should weigh these features against the added
% complexity of use.
% The package options above demonstrate how to enable PDF bookmarks
% (a type of table of contents viewable in Acrobat Reader) as well as
% PDF document information (title, subject, author and keywords) that is
% viewable in Acrobat reader's Document_Properties menu. PDF document
% information is also used extensively to automate the cataloging of PDF
% documents. The above set of options ensures that hyperlinks will not be
% colored in the text and thus will not be visible in the printed page,
% but will be active on "mouse over". USING COLORS OR OTHER HIGHLIGHTING
% OF HYPERLINKS CAN RESULT IN DOCUMENT REJECTION BY THE IEEE, especially if
% these appear on the "printed" page. IF IN DOUBT, ASK THE RELEVANT
% SUBMISSION EDITOR. You may need to add the option hypertexnames=false if
% you used duplicate equation numbers, etc., but this should not be needed
% in normal IEEE work.
% The latest version of hyperref and its documentation can be obtained at:
% http://www.ctan.org/pkg/hyperref

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{graphicx}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{times} 
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{
->, % makes the edges directed
>=stealth, % makes the arrow heads bold
node distance=2cm, % specifies the minimum distance between two nodes. Change if necessary.
every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
initial text=$ $, % sets the text that appears on the start arrow
}

\RequirePackage{todonotes}
\newcommand{\znjSide}[1]{\todo[color=orange!10]{\textbf{ZHAN:} #1}}
\newcommand\znj[1]{\textcolor{red}{#1}}
\newcommand{\parseInt}{\textsf{parseInt}}
\newcommand{\paexp}{$\textsf{PA}_{\exp}$}
\newcommand{\Def}{\hat{=}}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{resume}{Resume}
%\newtheorem{claim}{Claim}

\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
%\newtheorem{openq}{Open Question}
% \newcommand{\sextract}{\mathcal{S}}
% \newcommand{\ainput}{Input}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}

\author{Hao Wu,
        Yu-Fang Chen,
        Zhilin Wu, 
        Naijun Zhan% <-this % stops a space
\IEEEcompsocitemizethanks{
\IEEEcompsocthanksitem Academica Sinica.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://www.michaelshell.org/contact.html
\IEEEcompsocthanksitem xxx.}% <-this % stops a space
\thanks{xxxx.}
}

\author{\IEEEauthorblockN{Hao Wu\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Yu-Fang Chen\IEEEauthorrefmark{2}, 
Zhilin Wu\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Naijun Zhan\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}
\\
\IEEEauthorblockA{
\IEEEauthorrefmark{1} State Key Laboratory of Computer Science,\\
Institute of Software, Chinese Academy of Sciences, 100190, Beijing, China\\
Email: \{wuhao, wuzl, znj\}@ios.ac.cn}\\
\IEEEauthorblockA{\IEEEauthorrefmark{2}
Institute of Information Science, Academia Sinica, Taiwan\\
Email: yfc@iis.sinica.edu.tw}\\
\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Chinese Academy of Sciences,
Beijing, China}
}

% The paper headers
\markboth{FMCAD 2021}%
{Wu \MakeLowercase{\textit{et al.}}: A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

%\begin{document}
\maketitle{}

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion, which is an important function appearing in almost all scripting languages.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986. While the decision procedure by Point relies on quantifier eliminations and is expensive, we introduce various optimizations and provide an efficient implementation.
%    Besides the theoretical results, we also develop various optimizations and provide an implementation.
    We evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
%\keywords{ String-integer conversion \and Flat regular constraints \and Presburger arithmetic \and Exponential function \and Quantifier elimination.}

\end{abstract}

\iffalse     
    A Decision Procedure for String Constraints with the {\parseInt} Function and Flat Regular Constraints

    {\parseInt} converts strings to integers and is an important function in string manipulating programs. 
    Solving string constraints with {\parseInt} is challenging and undecidable in general. 
    The state-of-the-art string solvers resort to heuristics to solve such constraints. 
    In this work, we propose a decision procedure for a class of string constraints which 
    are boolean combinations of word equations, flat regular constraints, and {\parseInt} function. 
    The decision procedure is obtained by reducing the satisfiability problem to that of
    an extension of Presburger arithmetic with exponential functions, which is decidable. 
    We implement the decision procedure and carry out experiments to evaluate the performance. 
    The experimental results show the improvement against the state-of-the-art string solvers 
    in both precision and performance. To the best of our knowledge, this work represents the 
    first decision procedure for a class of string constraints with {\parseInt} function.
\fi

\section{Introduction}

solve string constraint is hard

strategy: do unsat and sat separately

strategy: using different procedure to (dis)prove validaity

for disprove validatity, there are two approches, first is bound string length

cannot handle $x.y \neq z  \wedge |x| > 2000$

more recent approach is flattening

it is known that word equaltion + flat regular constraints + len constraints
is decidable

it was unknown that whether word equations + flat regular constraints + len constraints + parseInt is decidable

\section{Preliminaries} \label{sec:pre}

In this section, we introduce some basic concepts and theories that will be used later. 
\subsection{Basic Concepts}

\paragraph{Sets and Strings}
We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the set of natural numbers and integers, respectively. 
$\mathbb{N}^+$ stands for the set of non-zero natural numbers. 
Let $\Sigma$ be a finite alphabet,
a string $w$ over $\Sigma$ is a sequence $a_1....a_n$ of characters from $\Sigma$.
Empty string is denoted by $\epsilon$.
$\Sigma^*$ denotes the set of all finite strings over $\Sigma$,
and $\Sigma_{\epsilon}$ stands for $\Sigma \cup \{\epsilon\}$.
For any string $w_1, w_2\in \Sigma^*$, 
we use $|w_1|$ to denote the length of $w_1$,
and $w_1\cdot w_2$ to denote the concatenation of $w_1$ and $w_2$.
A language $L$ over $\Sigma$ is a subset of $\Sigma^*$.

There are two types of variables in string constraints,
i.e., $X$, a set of string variables ranged over $\Sigma^*$,  
and $Z$,  a set of integer variables ranged over $\mathbb{Z}$.
As usual, an interpretation $I$ is a mapping from the set of variables $X\cup Z$ to the respective domain, 
essentially a pair of two mappings 
$I_X$ and $I_Z$, i.e., $I= (I_X, I_Z)$,  
where $I_X$ is a mapping in $X \mapsto \Sigma^*$ and $I_Z$ is a mapping in $Z \mapsto \mathbb{N}$.

\paragraph{Finite State Automata}
A Finite State Automaton is a tuple 
$\mathcal{A}=\langle Q,\Sigma,\Delta,$ $q_{\textit{init}},$ $q_{\textit{acc}}\rangle$, 
where $Q$ is a finite set of states, 
$\Sigma$ is the given alphabet,
$\Delta\subseteq Q\times \Sigma_\epsilon\times Q$ 
defines the transition relations in $\mathcal{A}$.
$q_{\textit{init}},q_{\textit{acc}}\in Q$ is the initial state and accepting state. 
A sequence $q_0 \langle a_1 \rangle  q_1 ... \langle a_n\rangle q_n$ is called accepting if $q_0 = q_{\textit{init}}$, $q_n = q_{\textit{acc}}$ and $q_{i-1}\langle a_i \rangle q_i \in \Delta$ for $1\le i\le n$.

\paragraph{Presburger Arithmetic} \label{PA}
The Presburger Arithmetic (PA) is a first order theory over signature 
$\Sigma_\mathbb{N}\Def  \{0,1,+,=\}$, where $0,1$ are constants, 
$+$ is a binary function and $=$ is a binary predicate.

PA can be axiomatized by the following axioms \cite{PA}

\begin{itemize}
    \item $\forall x, \neg (x+1=0)$
    \item $\forall x \forall y. x+1=y+1 \to x=y$
    \item $F(0) \wedge (\forall x. F(x)\to F(x+1)) \to \forall x. F(x)$ 
    \item $\forall x. x+0=x$
    \item $\forall x \forall y. x+(y+1)=(x+y)+1$
\end{itemize}

Given the domain $\mathbb{N}$,
the standard interpretation of PA interprets 
$0,1$ to $0_\mathbb{N},1_\mathbb{N}\in \mathbb{N}$
and $+,=$ to addition and equality over $\mathbb{N}$.
We call a PA formula without quantifiers a quantifier-free PA formula.

PA is a decidable theory, 
and the complexity of decidability is related to 
the number and locations of quantifiers.
Generally, 
the upper bound (on deterministic time and space) 
for deciding a formula of length $n$ is $2^{2^{2^{p n log(n)}}}$,
where $p>1$ is a constant\cite{Oppen69}.

\paragraph{Parikh Image}
Given an alphabet $\Sigma$ and a string $w\in \Sigma^*$, 
we define the set of Parikh variables 
$\Sigma^\# \Def \{a^\# \mid a\in \Sigma\}$.
The Parikh image of $w$ is a function 
$\mathbb{P}(w): \Sigma^\# \mapsto \mathbb{N}$,
which maps each symbol $a^\#\in \Sigma^\#$ to the number of occurrences
of $a$ in $w$.
For example, let $w\Def aabba$,
then $\mathbb{P}(w)(a^\#)=3,\mathbb{P}(w)(b^\#)=2$.

For a language $L\subseteq \Sigma^*$, 
define the Parikh image of $L$ to be 
$\mathbb{P}(L)\Def \{\mathbb{P}(w) | w\in L\}$.
We say a language $L$ is \emph{Parikh-definable} 
if $\mathbb{P}(L)$ can be characterized by a quantifier-free PA formula over $\Sigma^\#$, 
where $a^\#$ in the formula encodes the number of occurrences of $a$.
It is well known that
any context-free language (therefore regular language) 
is Parikh definable \cite{Parikh66}.

\paragraph{String Terms}
Given a finite alphabet $\Sigma$ and 
a finite set $X$ of string variables over $\Sigma^*$,
we define the set of terms $\textit{Terms}(\Sigma,X)$ 
to be the smallest set satisfying
\begin{itemize}
    \item[1] $\Sigma\cup \{\epsilon\} \cup X \subseteq \textit{Terms}(\Sigma,X)$;
    \item[2] if $t_1,t_2\in \textit{Terms}(\Sigma,X)$, then $t_1 \cdot t_2 \in \textit{Terms}(\Sigma,X)$.
\end{itemize}

We extend $I_X$ to $\textit{Terms}(\Sigma,X)$ by letting $I_X(\epsilon)=\epsilon$, 
for $a\in \Sigma, I_X(a)=a$,
and $I_X(t_1\cdot t_2)= I_X(t_1)\cdot I_X(t_2)$.

\paragraph{String Constraints} \label{par: string constraints}
Given a constraint $\phi$ and an interpretation $I$,
$I\models \phi$ denotes that $I$ satisfies $\phi$,
and $I$ is called a \emph{model} of $\phi$.
We use $\lVert \phi\rVert$ to denote the set of all models of $\phi$.

We define the following three forms of atomic string constraints:
\begin{itemize}
\item An equality constraint $\phi_e$ is of the form 
$t_1 = t_2$, where $t_1,t_2\in \textit{Terms}(\Sigma,X)$.
We define $\lVert \phi_e \rVert = \{I\mid I(t_1)=I(t_2)\}$.
Inequality constraints can be defined analogously.

\item A regular constraint $\phi_r$ is of the form 
$x\in L(\mathcal{A})$,
where $x\in \mathbb{X}$ and $\mathcal{A}$ is a finite state automaton.
We define $\lVert \phi_r \rVert = \{I\mid I(x)\in L(\mathcal{A})\}$.

\item A length constraint $\phi_l$ is a linear constraint over 
$Z \cup \{|x| \mid x\in X\}$, %the values of $|x|$ for all $x\in X$,
where $|\cdot |$ is the length function.
We define $\lVert \phi_l \rVert = \{I \mid I\models \phi_l \}$.
\end{itemize}

In Section~\ref{sec:string-solving},  
we will introduce a new form of atomic string constraints, i.e., \emph{string-number conversion constraints}.
A string constraint is a Boolean combination of atomic string constraints possibly with quantifications over $X\cup Z$. Other notions are same as in the first-order logic.

Giving a string constraint $\Psi$,
the problem of string constraints solving is to decide 
whether $\lVert \Psi \rVert$ is empty,
if not, to compute an interpretation $I$ that satisfies $\Psi$.

\subsection{Flat Automata and Flat Languages}

Given a string constraint $\Psi$,
the general problem of deciding whether $\lVert \Psi \rVert$ is empty is undecidable.
However, 
the problem becomes decidable when certain restrictions are imposed.
One of the restriction is by flat automata and flat languages, 
defined below.

\paragraph{Flat Languages and Automata}
For a fixed alphabet $\Sigma$,
we say a language $L$ over $\Sigma$ to be \emph{$\langle p,q \rangle$-flat} if 
there exist strings $w_1,...,w_q \in \Sigma^*$ such that
$|w_i|\le p$ for all $i:1\le i \le q$ 
and $L = (w_1)^*(w_2)^*...(w_q)^*$. 
We use $\alpha$ to denote $\langle p,q \rangle$, 
and call it the \emph{abstraction parameter} of $L$.
Intuitively,
a flat language with abstraction parameter $\alpha = \langle p,q \rangle$
consists of $q$ loops and the length of each loop body is equal or less than $p$.
For example,
$L = (ab)^*(a)^*(bb)^*$ is a $\langle 2,3 \rangle$-flat language.

Flat automata are a special form of finite state automata that 
recognize flat languages.
Fix the abstraction parameter $\alpha=\langle p,q\rangle$,
a flat automaton consists of $q$ loops,
each loop is a circle of $p$ states.
Formally, 
an $\alpha$-flat automaton contains $p q$ states at most,
and we name the states from $1$ to $p q$,
$1$ is the initial state and $(p q - p + 1)$ is the accepting state.
We use $\cdot$ as a placeholder for some symbol in $\Sigma_\epsilon$,
the transition relations of state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i \neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$,
    $(i, \cdot ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, then 
    $(i,\cdot , i-p+1) \in \Delta$;
    \item otherwise, $(i,\cdot, i+1) \in \Delta$.
\end{itemize}

A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$ is shown in 
figure (\ref{fig: FA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  of=4] (2) {$2$};
        \node[state, right of=4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, right of=3] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left,left] node{$a$} (2)
        (2) edge[bend left,right] node{$b$} (1)

        (3) edge[bend left,left] node{$a$} (4)
        (4) edge[bend left,right] node{$\epsilon$} (3)

        (5) edge[bend left,left] node{$b$} (6)
        (6) edge[bend left,right] node{$b$} (5);
    \end{tikzpicture}
    \caption{A $\langle 2,3 \rangle$-flat automaton 
that recognizes $L\Def  (ab)^*(a)^*(bb)^*$}
    \label{fig: FA}
\end{figure}

\paragraph{Generic Flat Languages and Automata}
Fix $\alpha = \langle p,q \rangle$,
we define the \emph{generic $\alpha$-flat language} is the union of all $\alpha$-flat languages, denoted by $\mathbb{F}(\alpha)$.
Now, we try to define an automaton that recognizes all $\alpha$-flat languages,
i.e., collects all behaviors of $\alpha$-flat automata.

Intuitively, 
the generic automaton is obtained by introducing a new alphabet (
a multi-set with $p q$ copies of the original alphabet) and 
adding more transitions (labels),
the states and the overall framework remain unchanged. 
In details, a generic $\alpha$-flat automaton is still a finite state automaton over
$\Sigma(\alpha)\Def \{a(i)| (a\in \Sigma_\epsilon) \wedge i\in \mathbb{N}:1\le i \le pq\}\cup \{\epsilon\}$.
The states are still named from $1$ to $pq$, 
the initial state is $1$ and the accepting state is $(pq-p+1)$.
The transition relations for state $i$ are defined as 
\begin{itemize}
    \item if $i\  \text{mod}\  p = 1$ and $i\neq pq-p+1$, then 
    $(i,\epsilon,i+p)\in \Delta$
    and $\forall s\in \Sigma_{\epsilon}. (i, s(i) ,i+1) \in \Delta$;
    \item if $i\  \text{mod}\  p = 0$, 
    $\forall s \in \Sigma_{\epsilon}. (i,s(i), i-p+1) \in \Delta$;
    \item otherwise, $\forall s \in \Sigma_{\epsilon}. (i,s(i), i+1) \in \Delta$.
\end{itemize}

For $\Sigma = \{a,b\}$, an example of generic $\langle 2,3 \rangle$-flat automaton is shown in figure (\ref{fig: GFA}).

\begin{figure}[ht]
    \centering 
    \begin{tikzpicture}
        \node[state,           ] (4) {$4$};
        \node[state, left  = 2cm of 4] (2) {$2$};
        \node[state, right = 2cm of 4] (6) {$6$};
        \node[state, initial, below of=2] (1) {$1$};
        \node[state, below of=4] (3) {$3$};
        \node[state, accepting, below of=6] (5) {$5$};

        \draw 
        (1) edge[above] node{$\epsilon$} (3)
        (3) edge[above] node{$\epsilon$} (5)

        (1) edge[bend left, pos =0.2 ,left] node{$a(1)$} (2)
        (1) edge[bend left, pos =0.5 ,left] node{$b(1)$} (2)
        (1) edge[bend left, pos =0.8 ,left] node{$\epsilon(1)$} (2)

        (2) edge[bend left, pos = 0.2 ,right] node{$\epsilon(2)$} (1)
        (2) edge[bend left, pos = 0.5 ,right] 
        node{$b(2)$} (1)
        (2) edge[bend left, pos = 0.8 ,right] 
        node{$a(2)$} (1)

        (3) edge[bend left, pos =0.2 ,left] node{$a(3)$} (4)
        (3) edge[bend left, pos =0.5 ,left] node{$b(3)$} (4)
        (3) edge[bend left, pos =0.8 ,left] node{$\epsilon(3)$} (4)

        (4) edge[bend left, pos = 0.2 ,right] node{$\epsilon(4)$} (3)
        (4) edge[bend left, pos = 0.5 ,right] 
        node{$b(4)$} (3)
        (4) edge[bend left, pos = 0.8 ,right] 
        node{$a(4)$} (3)

        (5) edge[bend left, pos =0.2 ,left] node{$a(5)$} (6)
        (5) edge[bend left, pos =0.5 ,left] node{$b(5)$} (6)
        (5) edge[bend left, pos =0.8 ,left] node{$\epsilon(5)$} (6)

        (6) edge[bend left, pos = 0.2 ,right] node{$\epsilon(6)$} (5)
        (6) edge[bend left, pos = 0.5 ,right] 
        node{$b(6)$} (5)
        (6) edge[bend left, pos = 0.8 ,right] 
        node{$a(6)$} (5);
    \end{tikzpicture}
    \caption{The generic $\langle 2,3 \rangle$-flat automaton}
    \label{fig: GFA}
\end{figure}

However,
the resulted automaton may accept languages that are not in $\mathbb{F}(\alpha)$,
because in different passes inside a loop, 
the automaton can choose different symbols between identical pairs. 
To avoid this problem,
we add a so-called purity condition on the accepting language of generic flat automata,
which is equivalent to intersecting the language of a generic flat automaton 
with a language that encodes the purity condition.

We say a string $w\in (\Sigma(\alpha))^*$ is pure if for all $i: 1\le i \le p q$,
and $a,b\in \Sigma$, 
$a\neq b \wedge \#w(a(i))>0$ implies $\#w(b(i))=0$.
Formally, the purity condition is defined by 
\begin{equation} \label{eq:purity}
 \bigwedge_{1\le i\le pq}\bigwedge_{a,b\in \Sigma, a\neq b} ({a(i)}^\#>0)\to ({b(i)}^\#=0)\, . 
\end{equation}

We denote the accepting language of the generic $\alpha$-flat automaton by $\mathbb{G}(\alpha)$.
Note that $\mathbb{G}(\alpha)$ is a language over $\Sigma_\alpha$,
but what we want is a language over $\Sigma$.
So we define a renaming function $R:\Sigma(\alpha)\mapsto \Sigma$ such that for all $a(i) \in \Sigma_\alpha, R(a(i))=a$,
and $R(\epsilon) = \epsilon$.
Define $\mathbb{G}'(\alpha) \, \Def \, \{R(w) \mid w\in \mathbb{G}(\alpha)\}$, 
for simplicity, we write
$\mathbb{G}'(\alpha)=R(\mathbb{G}(\alpha))$.

The important feature of generic flat autamata
is that every word $w\in \mathbb{G}(\alpha)$ is uniquely determined by its Parikh image $\mathbb{P}(w)$.

\subsection{Flattening}

The flattening technique was first introduced 
in \cite{Abdulla 2017}.
Fix an alphabet $\Sigma$ and an abstraction parameter $\alpha$, 
for a given atomic string constraint $\phi$, 
flattening $\phi$ with parameter $\alpha$ 
results in a new string constraint $\phi_\alpha$,
such that 
$R(\lVert \phi_\alpha \rVert) = \lVert \phi \rVert \cap \{I \mid \forall x\in X, I(x)\in \mathbb{G}'(\alpha)\}$,
where $R$ is the renaming function with its domain extended to interpretations in the normal manner.
Intuitively, it restricts $\phi$ to interpret 
string variables over $\mathbb{G}'(\alpha)$.

\cite{Abdulla 2017} discussed the flattening of basic 
string constraints including equality, integer, (regular) grammar and transducer constraints. 
For an atomic string constraint $\phi$,
the flattening $\phi_\alpha$ is still an atomic string constraint 
and is Parikh definable,
so its Parikh image can be expressed by a quantifier free PA formula.
Together with the purity condition,
we obtain an existential quantified PA formula $\rho$.
$\rho$ will be sent to a SMT solver,
if the solver returns an solution $\theta$,
then we can construct an interpretation for $\phi_\alpha$ from $\theta$,
otherwise it means $\phi$ is unsatistiable when 
string variables are interpreted  to $\alpha$-flat languages.

Take a regular constraint $\phi = x\in L(A)$ for example,
the flattening of $\phi$ resutls in a new finite state automaton $A'$ over $\Sigma(\alpha)$,
which encodes running $A$ ``in parallel" 
with the generic $\alpha$-flat automaton.
Let $\rho_1$ be the formula describing the Parikh image of $A'$,
which is a formula over variable sets $\Sigma(\alpha)^\#$.
Let $\rho_2$ be the purity condition \eqref{eq:purity}.
Then we obtain the PA formula $\exists (\Sigma(\alpha))^\#. \rho_1 \wedge \rho_2$.
In order to distinguish between different string variables,
we may replace $a(i)^\# \in \Sigma(\alpha)^\#$ by $(x,a(i))^\#$.

Since the structure of a flat automaton is decided 
by its abstraction parameter $\alpha$, 
a Counter-Example Guided Abstraction 
Refinement (CEGAR) framework is designed, which contains both an under- and an 
over-approximation module, to search the possible values of $\alpha$.
The termination for the overall algorithm is not guaranteed.

\section{Solving String Constraints with \textsf{parseInt} Function}\label{sec:string-solving}

The string-number conversion functions are commonly used functions
in most of programming languages,
for example,
\verb+parseInt()+ in Java and \verb+Int()+ in Python.
The functions usually take two parameters, 
a string over the agreed alphabet $\Sigma$
and an optional parameter denotes the radix.
They parse the string according to the rules indicated by the radix,
and return an integer denoted by the string.

From the view of string constraints,
string-number conversion functions give rise to a new form of string constraints
and are more expressive than length constraints.
So we consider extending string constraints with 
{\parseInt} function. 
As the general problem of string constraints is undecidable,
we still adopt the idea of flattening, 
i.e., variables are restricted to (generic) flat languages.
This problem has been investigated in \cite{POPL20},
which defined a special form of flat restriction (straight-line PFA) and 
proposed a heuristic search method.

In this section, 
we describe the problem of interest first, and then 
present an reduction from the problem of solving flat string constraints with {\parseInt} function
to the decidability problem of Presburger Arithmetic with exponential functions.
Hence, we identify a decidable subset of string constraints, which is the largest one with decidability so far to the best of our knowledge.

\subsection{String-Number Conversion Function}

Commonly, 
{\parseInt} function takes a string representation of a decimal
number and returns an integer.
Since we focus on the decidability,
we define a binary version of {\parseInt},
which takes a binary string and returns a decimal integer number.
For example,
${\parseInt}('111')=7$. 
Clearly,  our decision procedure given in this paper 
can be adapted to string constraints with other string to number conversion function without substantial change. 
Single quotation marks are used to distinguish a symbol like $'1'\in \Sigma$
from a number $1_\mathbb{N}\in \mathbb{N}$ when needed.

\begin{definition} Let $\Sigma_{\textit{num}}=\{'0','1'\}$,
 ${\parseInt}:\Sigma_{\textit{num}}^*\mapsto \mathbb{N}$ is recursively defined by:
    for $w\in \Sigma_{\textit{num}}^*$
    \begin{itemize}
        \item if $|w|=0$, i.e., $w=\epsilon$,  ${\parseInt}(w)=0$;
        \item if $|w|=1, {\parseInt}('w')=w_\mathbb{N}$;
        \item for $|w|\ge 2$ and $w=w_1 \cdot w_2$, 
        where $|w_1|, |w_2| \ge 1$, 
        ${\parseInt}(w) = {\parseInt}(w_1)2^{|w_2|}+{\parseInt}(w_2)$.
    \end{itemize} 
\end{definition}

Now, we introduce a new form of atomic string constraints: 
a {\parseInt} constraint $\phi$ is of the form 
$n \sim {\parseInt}(t)$,
where $n$ is an integer term, $\sim \in \{\le,<,=,>,\ge\}$ and $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$ is a string term.
$\lVert \phi\rVert \Def  \{I \mid I(n)\sim {\parseInt}(I(t))\}$.

In what follows, we only  consider the problem in the case when $t=x$ and $x$ is restricted to (generic) flat languages. 
%If $t$ is not a single variable, 
For the general case $t\in \textit{Terms}(\Sigma_{\textit{num}},X)$, 
it can be reduced to this special case by induction on the structure of $t$. 
%we can reduce the problem by separating $t$ (corresponding to the $|w|>2$ case in definition).

Given an $\alpha$-flat language $L$,
we assume $\alpha=\langle p,q \rangle$ and $L=(w_1)^*...(w_q)^*$,
where $p,q$ and $w_i(1\le i \le q) $ are known.
We further assume that $x = (w_1)^{\beta_1} ... (w_q)^{\beta_q}$,
then we have
\begin{align}
    &{\parseInt}(x) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_q)^{\beta_q}) \notag \\
    =&{\parseInt}((w_1)^{\beta_1} ... (w_{q-1})^{\beta_{q-1}})\cdot 2^{\beta_q |w_q|}
    + {\parseInt}((w_q)^{\beta_q}) \label{parse}
\end{align}
(\ref{parse}) is a recursive expression.
So we only need to deal with the basic case ${\parseInt}((w_q)^{\beta_q})$, 
where $w_q \neq \epsilon$
\begin{align}
    {\parseInt}((w_q)^{\beta_q}) &= \sum_{i=0}^{\beta_q-1} {\parseInt}(w_q)\cdot 2^{|w_q|\cdot i}  \notag \\
    &={\parseInt}(w_q)\frac{2^{|w_q|\cdot \beta_q}-1}{2^{|w_q|}-1}
    \label{parseInt-basic}
\end{align}
In (\ref{parseInt-basic}), since $w_q$ and $|w_q|$ are known, 
they can be regarded as constants.
The only unknown variable is $\beta_q$.

Combine (\ref{parse}) and (\ref{parseInt-basic}),
the constraint $n={\parseInt}(x)$ can be expressed by an arithmetic expression with 
$n$ and $(\beta_1,...,\beta_q)$,
inevitably with exponential components.

Take $n={\parseInt}((11)^a(10)^b)$ for example.
\begin{align}
    n=& {\parseInt}((11)^a)\cdot 2^{2b} + {\parseInt}((10)^b) \notag \\
    =& {\parseInt}(11)\cdot \frac{2^{2a}-1}{2^2-1}\cdot 2^{2b} + 
    {\parseInt}(10) \cdot \frac{2^{2b}-1}{2^2-1} \notag 
\end{align}

So we have 
\begin{equation}
    3\cdot n = 3\cdot 2^{2a+2b}-3\cdot 2^{2b}+2\cdot 2^{2b}-2. \notag
\end{equation}

Observe the form of the above equation,
$a,b,n$ are integer variables and either occur in 
an exponential term or a linear term.
This is always the case,
so the problem can be reduced to 
the decidability of PA with exponential function.

When $x$ of {\parseInt} constraints is restricted to the
generic $\alpha$-flat language ($\alpha$ is fixed),
the (\ref{parse}) and (\ref{parseInt-basic}) still hold 
but $w_i(1\le i\le q)$ is known.
However,
by definition,
the generic $\alpha$-flat language is the finite union of all $\alpha$-flat languages,
so we can enumerate all possible values for $w_i$.
In this way,
the problem can still be reduced to the decidability of PA with exponential function (\paexp), i.e.,
\begin{theorem} \label{thm:string-parInt}
If {\paexp} is decidable, then the satisfiability (validity) of string constraints with {\parseInt} in which all string variables 
are ranged over flat strings is decidable. 
\end{theorem}

\subsection{Decidability of \paexp} \label{sec:3.2}

For a first order theory $T$,
we say theory $T$ admits quantifier elimination (QE) if for any formula in $T$, 
there is a quantifier-free formula equivalent to it.
It is well-known that if a theory admits QE, 
then it is a decidable theory.

The formal definition of PA is given in section \ref{PA}.
Here we add the ordering predicate $\le$ into the signature,
which can be defined by $x \le y\, \Def \, \exists z. x+z=y$.
However, the theory PA so far does not admit QE, 
for example, consider the formula $\exists x.x = y + y$. 
We augment the theory with countable unary divisible predicates
$n|x$, where $n\in \mathbb{N}$, 
$n|x$ is true if and only if $x\ \text{mod}\ n=0$ holds.
This structure of PA that admits QE is denoted by $(\mathbb{N},+)$.

We then introduce a theory {\paexp}, denoted by $(\mathbb{N},+,2^x)$, that we work on.
\begin{definition}
    Let $\mathcal{L}=\{0,1,+,\le,n \, |x(n\in \mathbb{N}),2^x,l_2(x)\}$, 
     $(\mathbb{N},+,2^x)$ be a $\mathcal{L}$-theory that has domain $\mathbb{N}$, where 
    \begin{itemize}
        \item  $2^x$ is interpreted to the exponential function of $2$ over $\mathbb{N}$; 
        \item interpretations of $0,1,+,\le,=$ are consistent with PA;
        \item for $n\ge 1$,$n|x$ holds iff $\exists y.x=ny$;
        \item $2^0=1$, for $n \ge 1, 2^n = 2^{n-1}+2^{n-1}$;
        We further assume that if $m,n\in \mathbb{N}, m\le n$, 
        then $2^{m-n} = 1$
        \item $l_2(0)=0$; for $n \ge 1,l_2(n) = y$ iff $2^y \le n < 2^{y+1}$;  $l_2(m-n) = 0$
        if $m,n\in \mathbb{N}$ and $m\le n$. 
    \end{itemize}
\end{definition}

$\lambda_2(x) = 2^{l_2(x)}$ can be defined by $l_2(x)$,
intuitively, $\lambda_2(x)$ means the maximal power of 2 that is not larger than $x$.
Then we have $\lambda_2(x) \le x \le 2\lambda_2(x)-1$,
which will be useful in our proof.

\begin{definition}
    For a strictly increasing function $f:\mathbb{N}\mapsto \mathbb{N}$, 
    $f$ is said to be compatible with addition 
    if for every $m\in \mathbb{N}^+$,  $f$ modulo $m$ is periodic, 
    and for any term $A(x)=\sum_{1\le i\le n} a_i f(x+b_i)$, 
    where $n \in \mathbb{N}^+, a_i ,b_i \in \mathbb{Z}$,
    one of the following holds:
    \begin{itemize}
        \item $A(x)$ is bounded; 
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. A(x+\Delta_A)\ge f(x)$;
        \item there exists a constant $\Delta_A\in \mathbb{N}^+$ such that 
        $\forall x. -A(x+\Delta_A)\ge f(x)$.
    \end{itemize}
\end{definition}

%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
by \cite{Semenov84}.
Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Point's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

In the next section, within Point's framework, we give a revised algorithm
 to the decision problem of {\paexp} with 
both subtle improvements and crucial corrections, that 
provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
Theorem~\ref{thm:string-parInt}.

\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
that is, $\exists x.\theta(x,\bar{y})$. 
For eliminating quantifiers in arbitrary formula, 
we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.

We divide the whole QE algorithm into 4 steps.
The first step $\textbf{Normalization}$ can be viewed as a pre-processing step, which substitutes ``complex" terms like $2^{2^x}$, $l_2(3x+y)$ with ``simple" terms by introducing new variables $x_i$. Now we move forward to handle a ``simpler" formula, however, at the cost of more quantified variables.

The other three steps undertake the task to eliminate the introduced variables $x_i$ and the original variable $x$ (will be denoted by $x_0$) one by one. 
The main body of \textbf{QE-with-Order} step is a loop to enumerate all possible orders among quantified variables. 
In each iteration, according the given (decreasing) order (corresponding to a for loop), we invoke \textbf{QE-exp} or 
\textbf{QE-linear} to eliminate these variables one by one. 
%depending on  eliminates the maximal variable in the remaining $\bar{x}$. So it is necessary to specify an ordering for $\bar{x}$ at the beginning, as we will see, this is the main reason for the complexity of this algorithm.
During each iteration of the inner loop (the for loop), 
if the maximal $x_i$ in $\bar{x}$ occurs in an exponential term in an atomic formula, 
we will invoke \textbf{QE-exp} to produce a formula equivalent to the atom where $x_i$ occurs linearly;  
otherwise, if $x_i$ occurs linearly in the formula, \textbf{QE-linear} will be invoked to eliminate all occurrences of $x_i$, 
and this procedure is similar to the classic QE algorithm for PA.

A more detailed description is given below.

\subsection{Normalization}

In order to show $(\mathbb{N},+,2^x)$ admits QE , 
it is sufficient to show that any 1-existential formula $\exists x.\theta(x,\bar{y})$ indeed does, 
where $\theta(x,\bar{y})$ is a quantifier-free formula. 
However, the form of $\exists x.\theta(x,\bar{y})$ is unknown so it may contain terms difficult to handle such as $2^{3x+y+1}$ or $l_2(x)$.
The \textbf{Normalization} step simplifies these terms by introducing new variables, 
for example, $\exists x.2^{3x+y+1}>10$ is equivalent to
$$\exists x\exists x_1.2^{x_1}>10 \wedge x_1 =3x+y+1\,.$$

Logarithm functions are replaced by exponential functions, take $\exists x.l_2(x)>3$ for example, $l_2(x)$ is replaced by a new variable $x_1$, and we have  
$$\exists x\exists x_1. x_1 > 3 \wedge 2^{x_1}\le x \wedge x\le 2^{x_1+1}-1\,.$$

The \textbf{Normalization} step goes like this, first transform the given formula $\theta(x,\bar{y})$ into the Negation Normal Form (NNF).
Then $\theta(x,\bar{y})$ becomes a Boolean combination (with only $\wedge$ and $\vee$) of literals.

Make substitutions by introducing new variables according to the \textit{while} statements in the pseudo-code.For consistency, we rename the original $x$ to be $x_0$ and assume $n$ new quantified variables are introduced.

After introducing new variables and substitution, 
we obtain a formula where if an exponential term contains a quantified variable $x_i(0\le i\le n)$, the term should be $2^{x_i}$.
We then collect all terms with $x_i(0\le i\le n)$ together, 
it will be of the form $s(\bar{x})\Def \sum_{i=0}^{n} a_i 2^{x_i} + \sum_{i=0}^n b_i x_i$, where $a_i,b_i(0\le i \le n)$ are all constants.
Other terms including constants and terms of $\bar{y}$ are collected,  
denoted by $t(\bar{y})$. 
Since $\bar{y}$ are free variables, 
we will regard $t(\bar{y})$ as a constant.
Inequalities and equalities will all be transformed into $s(\bar{x})\le t(\bar{y})$, for example, $s(\bar{x}) = t(\bar{y})$ will be replaced by $s(\bar{x})\le t(\bar{y})\wedge t(\bar{y})\le s(\bar{x})$.

At the end, the resulted formula $\theta'$ only contains literals of the forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ and $\neg  (k|s(\bar{x})+t(\bar{y}))$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{1-existential formula $\exists x.\theta(x,\bar{y})$}
    \KwOut{n-existential formula $\exists \bar{x}.\theta'(\bar{x},\bar{y})$}

    $\theta' :=  \theta(x,\bar{y})$ \;
    Transform $\theta'$ into NNF\;
    \tcp{ i is used for counting the introduced variables }
    i = 1\; 
    \tcp{ $\bar{y}$ are regarded as constants} 
    \While{there is a term $l_2(t)$, $t$ is not a constant}
    {
        $\theta' :=  \theta'[x_i/l_2(t)] \wedge (2^{x_i}\le t) \wedge (t\le 2^{x_i+1}-1)$\;
        $i:= i+1$\;
    }
    \While{there is a term $2^t$, $t$ is not a variable $x_j(j<i)$ or a constant}
    {
        $\theta' :=  \theta'[2^{x_i}/2^t]\wedge (x_i\le t) \wedge (t \le x_i)$\;
        $i:= i+1$\;
    }
    $n:= i-1$\;
    \tcp{ Collect quantified variables $x_i$}
    Transform all atoms into forms $s(\bar{x})\le t(\bar{y})$, $k|s(\bar{x})+t(\bar{y})$ or $\neg  (k|s(\bar{x})+t(\bar{y}))$\;
    Return $\exists x_0,...,\exists x_n. \theta'$

    \caption{Normalization}
\end{algorithm}

\subsection{QE-with-order} 
During the \textbf{Normalization} step, 
we simplify the origin formula at the cost of more quantified variables. 
Suppose we get $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
and it has $n$ new variables $x_i$ ($x$ is denoted by $x_0$).
As a consequence, we need to eliminate all quantified variables one by one.

To the end, we denote the set of all orders among the $n+1$ variables by 
 $S_{n+1}$, and then enumerate these orders one by one in the outer \textit{for} loop. 
 For the considered order $\sigma$, we first add the ordering information to 
 the quantifier free formula, and then according to the 
 order, to eliminate $x_{\sigma(i)}$ for $i=n$ to $0$ by invoking \textbf{QE-exp} first, and then invoking \textbf{QE-linear} (the inner \textit{for} loop). 
 In each iteration of the inner \textit{for} loop, 
if $x_{\sigma{(i)}}$ occurs in an exponential term in an atomic formula, 
\textbf{QE-exp} is invoked and a formula   equivalent to the atom is returned, in which $x_i$ occurs linearly;  
then $x_i$ occurs linearly in the formula, thus \textbf{QE-linear} is further invoked to eliminate all occurrences of $x_i$. 
The returned formula is the disjunction of all formulas resulted in all iterations of the outer \textit{for} loop.

\iffalse
put , We now specify an order of $\bar{x}$, and recursively eliminate the maximal element in the remaining $\bar{x}$. 
That is to say,
every execution of \textbf{QE-exp} and \textbf{QE-linear} will eliminate the largest element $x_i$ in $\bar{x}$.
However, as you may notice, if there is no clue about the order of $\bar{x}$, 
we will have to use the idea of permutation group and the number of sub-cases will blow up, in the worst case, $(n+1)!$ cases.

Here is how \textbf{QE-exp} and \textbf{QE-linear} are invoked in
\textbf{Ordering and QE}. 
\fi

%我们这里还是考虑用循环群来做，因为如果每一次只找到最大的$x_i$，那么递归的执行还是需要整个$\bar{x}$的顺序。但如果一开始就知道了，后面有些操作可以简化

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{a normalized $(n+1)$-existential formula $\exists \bar{x}.\theta(\bar{x},\bar{y})$,
    where $\bar{x}=(x_0,...,x_n)$} 
    \KwOut{a equivalent quantifier free formula without $\bar{x}$}

    Let $S_{n+1}$ denote the group of permutations on $\{0,1,...,n\}$\;
    $\phi :=  \textit{False}$\;
    \For{each $\sigma \in S_{n+1}$ }
    {
        $\theta_{\sigma}:=  \theta(\bar{x},\bar{y})\wedge \bigwedge_{j=0}^{n-1}(x_{\sigma(j)}\le x_{\sigma(j+1)})$\;

        \tcp{recursively eliminate the maximal $x_i$ in $\bar{x}$}
        %\znj{How to effectively determine the maximal $x_i$?} 
        \For{i from n to 0}
        {
            \tcp{if $x_{\sigma(i)}$ occurs exponentially in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-exp}(x_{\sigma(i)},\theta_{\sigma})$\;
            \tcp{now $x_{\sigma(i)}$ occurs only linearly in $\theta_\sigma$}
            $\theta_{\sigma}:= \textbf{QE-linear}(x_{\sigma(i)},\theta_{\sigma})$\;
        }
        $\phi :=  \phi \vee \theta_{\sigma}$\;
    }
\iffalse
    \For{i from $0$ to $n$}
    {
        \tcp{ specify the maximal quantified variable}
        $\theta_i :=  \theta(\bar{x},\bar{y}) \wedge (\bigwedge_{0\le j \le n} x_i - x_j \ge 0)$\;
        \tcp{ eliminate $x_i$}
        \While{$x_i$ occurs in an exponential term in formula $\tau(x_i,\bar{x},\bar{y})$}{
            \tcp{ QE-exp outputs a formula equivalent with $\tau(x_i,\bar{x},\bar{y})$ where $x_i$ occurs linearly }
            $\theta_i :=  \theta_i[\text{QE-exp}(\tau(x_i,\bar{x},\bar{y}))/\tau(x_i,\bar{x},\bar{y})]$\;
        }   
        transform $\theta_i$ into disjunction normal form $\theta_i = \bigvee_s \theta_{i,s}$\;
        \For{ every $\theta_{i,s}$ }{
            $\theta'_{i,s} :=  \text{QE-linear}(x_i,\theta_{i,s})$
        }
        $\theta'_i =\bigvee_s \theta'_{i,s}$\;
    }
\fi
    output $\phi$
    \caption{QE-with-order}
\end{algorithm}
%\znj{I cannot understand why you need to do QE for all permutation $\sigma$? I think 
%one is enough.}

\subsection{QE-exp}

\textbf{QE-exp} and \textbf{QE-linear} are the most technical parts of our QE algorithm.
As mentioned, 
\textbf{QE-exp} takes $\theta_{\sigma}(\bar{x},\bar{y})$ and  variable $x_{\sigma(i)}$ (the maximal variable among all quantified variables that are not eliminated yet according to $\sigma$) as 
inputs, and  outputs an equivalent formula in which $x_{\sigma(i)}$ occurs linearly.
The ideal case is that the formula $\theta(\bar{x},\bar{y})$ itself contains no $2^{x_{\sigma(i)}}$ terms, so we can omit this step and directly go to \textbf{QE-linear}.
For simplicity, we will abuse   $x_i$ for $x_{\sigma(i)}$, 
and  $\bar{x}$ for $(x_{\sigma(0)},...,x_{\sigma(i-1)})$ ( remind that $x_{\sigma(i+1)},...,x_n$ have been eliminated already).

After normalization, the formula $\theta(\bar{x},\bar{y})$ contains atoms of three forms corresponding to 
the predicates $\le$, $|$ and negation of $|$, respectively.
The problem will be discussed in 2 cases depending on the form of the atom $\tau$ that contains $2^{x_i}$, 
corresponding to \textbf{QE-exp-ineq} or \textbf{QE-exp-div}.

We first give the whole procedure of \textbf{QE-exp},
and then provide the details of the two sub-routines.
\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i$ and $\theta$, $x_i$ is larger than $x_0$ to $x_{i-1}$
    and occurs exponentially in $\theta$} 
    \KwOut{a formula equivalent to $\theta$ where $x_i$ occurs linearly}

    \While{there is an atom $\tau(x_i,\bar{x},\bar{y})$ contains $2^{x_i}$}
    {
        $\Psi:= \textit{False}$\;
        \eIf{$\tau$ is of the form $a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$}
        { 
            \tcp{ \textbf{QE-exp-ineq} case}
            $A :=  \sum_{j=0}^{i-1}|a_j|$, $B:= b_i + \sum_{j=0}^{i-1}|b_j|$\;
            $B':= 2(l_2(B)+3)$, $\delta:=  l_2(A)+3$\;
            If $a_i>0$ then $\alpha := l_2(t(y))-l_2(a_i)$
            else $\alpha := l_2(-t(y))-l_2(-a_i)$\;

            $\rho_1 :=  (x_i \le \alpha -1 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee ( x_i \le \alpha -1 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta} (x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))$\;

            $\rho_2 :=  (x_i = \alpha) \wedge \tau[\alpha/x_i]$\;
            $\rho_3 :=  (x_i = \alpha+1) \wedge \tau[\alpha+1/x_i]$\;
            $\rho_4 :=  (x_i \ge \alpha+2 \wedge \bigvee_{0\le k\le B'}(x_i=k \wedge \tau[k/x_i] ))$

            $\quad \vee (x_i \ge \alpha +2 \wedge x_i\ge B' \wedge \bigvee_{0\le k \le \delta}(x_i = x_{i-1}+k \wedge \tau[x_{i-1}+k/x_i]))
            $\;
            \eIf{$a_i>0$}
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \le \alpha -1 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$
            }
            {
                $\Psi := \rho_1 \vee \rho_2 \vee \rho_3 \vee \rho_4 
                \vee [x_i \ge \alpha +2 \wedge x_i \ge B' \wedge x_i \ge x_{i-1}+\delta]$    
            }
            }
        {
            \tcp{ \textbf{QE-exp-div} case, the atom is of the form $d|t(x_i,\bar{x},\bar{y})$}
            \tcp{ let $d=2^r d_0$ where $d_0$ is odd}
            $\rho_5 := \bigvee_{p=0}^{r-1} [\tau(p,\bar{x},\bar{y})\wedge x=p]$\;
            $\rho_6 := \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
            +  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r]
            $\;
            $\Psi :=  \rho_5 \vee \rho_6$
        }
        replace $\tau$ by $\Psi$ in $\theta$\;
    }
    \caption{QE-exp}
\end{algorithm}

\subsubsection{QE-exp-ineq}

This algorithm deals with the case where $\tau$ is an inequality of the form 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$

We now try to eliminate $2^{x_i}$ in $\tau$, the idea is to find a bound for $x_i$, 
either by constants or by other variables.
We will prove the following theorem.

\begin{theorem} \label{thm:exp-ineq}
Assume that $x_i$ is the maximal variable in $\bar{x}$ according to the given order. 
Given an inequality $\tau(x_i,\bar{x},\bar{y})$ of the form %, suppose $x_i$ occurs in an exponential term, that is 
$$\tau(x_i,\bar{x},\bar{y}) \Def  a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j2^{d x_j} + \sum_{k=0}^{i}b_k x_k \le t(\bar{y})$$
with $a_i \neq 0$, 
let $A\Def \sum_{j=0}^{i-1}|a_j|$, 
$B\Def |b_i| + \sum_{j=0}^{i-1}|b_j|$, 
$B'\Def 2(l_2(B)+3)$,
$\delta\Def  l_2(A)+3$, then 
\begin{itemize}
    \item if $a_i > 0$, let $\alpha \Def l_2(t(y))-l_2(a_i)$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
    \end{itemize}
    \item if $a_i < 0$, let $\alpha \Def l_2(-t(y))-l_2(-(a_i))$
    \begin{itemize}
        \item if $x_i \le \alpha -1$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ \textbf{does not} hold.
        \item if $x_i \ge \alpha +2$, $x_i \ge B'$ and $x_{i} \ge x_{i-1} +\delta $, then $\tau(x_i,\bar{x},\bar{y})$ holds.
    \end{itemize}
\end{itemize}
\end{theorem}

Before proving \textbf{Theorem 1}, we need  the following lemma to estimate linear terms.

\begin{lemma} \label{lem:1}
For any $n,m\in \mathbb{N}$, if $n\ge m\ge 1$ and $x\ge 2(l_2(n)-l_2(m)+1)$, then 
$nx\le m2^x$ holds.
\end{lemma}

\begin{proof}
First we can prove that for any $N\in \mathbb{N}$, $x\ge 2N \implies 2^N x\le 2^x$. Let $N\Def l_2(n)-l_2(m)+1$, then we have $x\ge  2N \implies n x\le 2\lambda_2(n) x\le 2^N \lambda_2(m) x \le m 2^x $. \qed 
\end{proof}

Then we give the proof for Theorem~\ref{thm:exp-ineq}.
\begin{proof}
We only prove for the $a_i > 0$ case, the other case is analogous. 
The goal is to find a bound for $x_i$ such that 
the values of the atoms containing $x_i$ keep constant when $x_i$ is greater than the bound.

Note that $x_i$ is the largest among $\bar{x}$. 
suppose $x_i > x_{i-1} + \delta$, and let 
 $\delta \Def  l_2(A)+3$, then 
 \begin{equation} 
   2^{-\delta}A = \frac{A}{8\lambda_2(A)}\le \frac{1}{4}.   \label{eq:thm-ineq-1}
 \end{equation}
 When $x_i \ge B' = 2(l_2(4B)-l_2(1)+1)$, according to Lemma~\ref{lem:1}, 
\begin{equation} 
    4 B x_i \le 2^{x_i}. \label{eq:thm-ineq-2}
\end{equation}

When $x_i \ge \alpha+2$,
\begin{align}
    a_i2^{x_i}+ \sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k 
    & \ge a_i 2^{x_i} - 2^{x_i -\delta}A  - Bx_i & \notag \\
     & \ge 2^{x_i}(a_i -\frac{1}{4} -\frac{1}{4}) &  (\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}})\notag \\
    %= &(a_i-\frac{1}{2})2^{x_i} \notag \\
    %\ge &\frac{a_i}{2}2^{x_i} \notag \\
    & \ge  \frac{a_i}{2}\frac{4\lambda_2(t(\bar{y}))}{\lambda_2(a_i)} & 
      (\mbox{ Def. of } \alpha) & \notag \\
     & \ge  t(\bar{y}) & \notag 
\end{align}
So we conclude that %when $x_i \ge \alpha+2,x_i\ge B',x_i \ge x_{i-1}+\delta$,
$\tau(x_i,\bar{x},\bar{y})$ keeps \textit{false} in this case.

%\znj{I think there are some mistakes in the above proof, pls check it!}
When $x_i \le \alpha-1$, similarly we have
\begin{align}
    \tau(x_i,\bar{x},\bar{y})\notag 
    & \le  a_i 2^{x_i} + 2^{x_i -\delta}A  + Bx_i &  \notag \\
    & \le  2^{x_i}(a_i +\frac{1}{4} +\frac{1}{4}) &(\mbox{by \eqref{eq:thm-ineq-1} and  \eqref{eq:thm-ineq-2}}) \notag \\
   & \le  2\lambda_2(a_i)\frac{\lambda_2(t(\bar{y}))}{2\lambda_2(a_i)} &  (\mbox{ Def. of } \alpha) \notag \\
   &  \le  t(\bar{y}) & \notag 
\end{align}
which indicates that when  $x_i \le \alpha-1$, $\tau(x_i,\bar{x},\bar{y})$ keeps \textit{true}. \qed 
\end{proof}

\subsubsection{QE-exp-div}

If $x_i$ occurs exponentially in an atom $\tau(x_i,\bar{x},\bar{y})$ of the form
$$\tau(x_i,\bar{x},\bar{y}) \Def  d| a_i 2^{x_i}+\sum_{j=0}^{i-1} a_j 2^{x_j} + \sum_{k=0}^{i} b_k x_k+t(\bar{y}) \quad (a_i \neq 0)\,,$$
the algorithm \textbf{QE-exp-div} outputs an equivalent formula without $2^{x_i}$ terms.
The idea is that $a_i 2^x$ modulo $d$ is a periodic function when $x$ is large enough and the period can be computed.

Let $d = 2^rd_0$, $d_0$ is an odd natural number. 
According to Euler's Theorem, 
$gcd(2,d_0)=1$ implies
$2^{\phi(d_0)} \text{mod}\ d_0 = 1$,
where $\phi$ is the Euler function.
Consider function $f(n)=(2^n\ \text{mod} \ d)$,
when $n\ge r$, 
$f(n)$ becomes a periodic function and its period is a divisor of $\phi(d_0)$ because
\begin{align}
    f(n+\phi(d_0)) 
    &=2^{n+\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot 2^{\phi(d_0)}\ \text{mod}\ d &\notag\\
    &=2^n\cdot (k d_0 + 1) \ \text{mod}\ d &
    (\mbox{assume } 2^{\phi(d_0)}=k d_0 +1) \notag \\
    &= 2^n \ \text{mod}\ d &
    (\mbox{when } n\ge r, 2^n d_0\ \text{mod}\ d = 0 )\notag\\
    &= f(n) \notag
\end{align}

When $x_i\le r-1$, we just enumerate all possible value of $x_i$, i.e., 
$$\rho_5 \Def \bigvee_{p=0}^{r-1} \tau(p,\bar{x},\bar{y})\,.$$
When $x_i\ge r$, $\tau(x_i,\bar{x},\bar{y})$ is equivalent to
$$
\rho_6 \Def \bigvee_{q=0}^{\phi(d_0)-1} [d|(a_i 2^{r+q}+\sum_{j=0}^{i-1} a_j 2^{x_j}
+  \sum_{k=0}^{i} b_k x_k+t(\bar{y})) \wedge \phi(d_0)|(x_i-r-q) \wedge x_i \ge r] \,.
$$
            Therefore,  $\tau(x_i,\bar{x},\bar{y})$ is equivalent to  $\rho_5 \vee \rho_6$.

%\znj{\textbf{Solved}.More explanation or a formal proof is needed. }

\subsection{QE-linear}

After \textbf{QE-exp}, 
we obtain a formula (still denoted by $\theta(x_i,\bar{x},\bar{y})$) without $2^{x_i}$ terms,
i.e., $x_i$ occurs only linearly.
Now we wish to construct a formula without $x_i$ equivalent to $\exists x_i.\theta(x_i,\bar{x},\bar{y})$.
This can be done by 
following Cooper's QE algorithm for PA,
treating all $x_j (j<i)$ as free variables (like $\bar{y}$).
The procedure contains 3 steps.

Step 1: Put $\theta(x_i,\bar{x},\bar{y})$ in NNF form, replace atomic formulae containing symbols other than $\le,|$ by equivalent formulae only with $\le$, that is, $x=a$ by $x\le a \wedge a\le x$, $x< a$ by $x\le a+1$, $x\neq a$ by $x>a \vee x<a$,  $x\geq a$ by $-x \le -a$, and $x>a$ by $-x \le -a-1$.
In inequality atoms, collect terms of $x_i$
to one side and guarantee the coefficients of $x_i$ are positive.

Step 2: Let $d$ be the least common multiple of all coefficients of $x_i$. 
For atoms with $x_i$, multiply all terms by a factor so that the coefficient of $x_i$ is $d$.
We introduce a fresh variable $x_i'$, replace all occurrence of $dx_i$ by $x_i'$,
and denote the resulted formula by $\theta'(x_i')$.
Note that $x_i'$ is a multiple of $d$, 
so we set $\theta'=\theta'\wedge d|x'$.

We classify all atoms in $\theta$ into the following three sets:  
$L$ denotes all atoms of the form $t_l(\bar{x},\bar{y})\le x_i'$,
$U$ denotes all atoms of the form $x_i'\le t_u(\bar{x},\bar{y})$,
$M$ denotes all atoms of the form $k_m|x_i'+t_m(\bar{x},\bar{y})$ and their negations,
where $t(\bar{x},\bar{y})$ with a subscript is any term of $\bar{x}$ and $\bar{y}$ and 
$k_m \in \mathbb{N}$ for $m\in M$. % are integers in $$. 
%Then all atoms with $x_i'$ are of one of the types $L,U,M$.

Step 3: Let $\delta$ be the least common multiple of $\{k_m \mid m\in M\}$.
Construct $\theta'_{+\infty}(x_i')$ from $\theta'(x_i')$ by  
 replacing $\textit{true}$ for all atoms in $L$,
and $\textit{false}$ for all atoms in  $U$.
%We use $\theta'[t]$ to denote replacing all $x_i'$ by a term $t$. 
Then $\exists x_i'.\theta'(x_i')$
is equivalent to

$$\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee 
\bigvee_{j=0}^{\delta-1} \bigvee_{u\in U} \theta'[t_u(\bar{x},\bar{y})-j/x_i']$$

The first disjunction $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i']$ corresponds to the case where $x_i'$ is large enough 
so that all atoms in $L$ are \textit{true} and all atoms in  $U$ are \textit{false}. Thus $\theta'_{+\infty}$  contains only divisibility literals.
If there is a number $n (0\le n \le \delta-1)$ such that $ \theta'_{+\infty}[n/x_i']$ is evaluated to be \textit{true},
then for every $\lambda\in \mathbb{N}$,
$\theta'_{+\infty}[n+\lambda\delta/x_i']$ is also evaluated to be \textit{true}.
Hence, there exists $x_i'$ large enough to satisfy $\theta'$.

The second disjunction corresponds to the case that for 
some $u\in U, x_i'\le t_u(\bar{x},\bar{y})$ holds.
In this case, 
select the minimal $t_u(\bar{x},\bar{y})$
from atoms in $U$ that holds,
then there exists a solution for $x_i'$ such that $x_i'$ is in the interval $[t_u(\bar{x},\bar{y})-\delta+1,t_u(\bar{x},\bar{y})]$.

Here we describe Cooper's algorithm in pseudo code.
We will use $\theta\{\tau'/\tau\}$ to denote substitute all atoms $\tau$ with $\tau'$,
distinguished from term substitution $\theta[t'/t]$.

\begin{algorithm}[t]
    \SetAlgoLined
    \KwIn{$x_i,\theta(x_i,\bar{x},\bar{y})$,
    where $x_i$ occurs linearly} 
    \KwOut{A equivalent formula without $x_i$}

    Collect terms of $x_i$ in each atom\;
    Let $d$ to be the least common multiple of all coefficients of $x_i$\;
    Multiply each atom by a factor so that coefficient of $x_i $ is $d$\;
    Let $\theta'(x_i'):= \theta[x_i'/dx_i]$\;
    $\theta'(x_i') := \theta'(x_i') \wedge d| x_i'$\;
    \tcp{ Now the atoms have 3 different forms}
    \tcp{ $t_l(\bar{x},\bar{y})\le x_i'$, $x_i' \le t_u(\bar{x},\bar{y})$, $k_m|x_i'+t_m(\bar{x},\bar{y})$(or its negation)}
    \tcp{ where $l\in L,u\in U,m\in M$ are the index sets}
    $\delta :=  lcm\{k_m|m\in M\}$\;
    \For{all atom $\tau_l \in L$}
    {
        $\theta'_{+\infty}:=  \theta'\{\textit{true}/\tau_l\}$
    }
    \For{all atom $\tau_u \in U$}
    {
        $\theta'_{+\infty}:=  \theta'_{+\infty}\{\textit{false}/\tau_u\}$
    }
    output $\bigvee_{j=0}^{\delta-1} \theta'_{+\infty}[j/x_i'] \vee \bigvee_{j=0}^{\delta-1} \bigvee_{u\in U}\theta'[t_u(\bar{x},\bar{y})-j/x_i']$
    \caption{QE-linear (Cooper's QE algorithm for PA)}
\end{algorithm}

%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
% 之前的算法

\subsection{Compared with Francoise's Origin Algorithm}

Francoise's algorithm \cite{Francoise} has four steps,
corresponding to the four steps of our algorithm respectively.
So we will use the names of our steps to refer to them. 
The divisible predicate $n|x$ is not included in his theory,
instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
with some minor differences due to the setting of the theory.
The main differences between the two algorithms lies in the following aspects.

In \textbf{QE-exp},
we adopt the same idea to find sufficient conditions
and necessary conditions for the formula to holds.
But our strategy for choosing parameters are easier to understand and use.
A flaw lies in the $a_i<0$ case in Francoise's algorithm 
\footnote{
In the $a_i<0$ case of \textbf{QE-exp-ineq},
Francoise's algorithm assumes falsely $t(y)\ge 0$.
The reason for this mistake might be the confusion of 
two subtraction operators.
The example mentioned in \ref{sec:3.2} shows this case.
},
our \textbf{QE-exp-ineq} corrects this flaw by 
treating $a_i<0$ case similarly to the $a_i>0$ case.

\textbf{QE-linear} in Francoise's algorithm in some sense includes our
\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
In this step,
Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
We instead invoke the QE algorithm of PA without referring to permutation groups, 
other improved QE algorithm can also be used to simplify this step.

Francoise's algorithm has more restrictions on the form
of the formula.
Before \textbf{QE-linear} and \textbf{QE-exp} step,
it needs to transform the formula into disjunction normal form (DNF)
and deals with the basic case when input formula $\theta$ is 
a conjunction of atoms. 
It is well known that transforming an arbitrary formula into 
DNF is costly and the length of the formula may increase exponentially.
So our algorithm remove these limitations and assume no special forms of
the formula other than NNF.

\section{Complexity Analysis and Case Study}
In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
\subsection{Complexity Analysis}

In this part, we estimate the complexity of our algorithm.
It can be divided into two part, 
the first part convert a string constraint $\phi$ to a PA formula with exponential functions,
and the second part is the QE procedure on the obtained formula.

Given a finite alphabet $\Sigma$ and $\alpha=\langle p,q \rangle$, the size of $\Sigma$ is denoted by
$|\Sigma|$. 
The generic $\alpha$-flat language contains $O(|\Sigma|^{pq})$ $\alpha$-flat languages

\paragraph{Complexity for QE}
In \textbf{QE-with-order},
it reduces the origin problem to $(n+1)!$ sub-cases,
and for each case,
we recursively invoke \textbf{QE-exp} and \textbf{QE-linear}.
So the key is to analysis the inner \textit{for}
loop in \textbf{QE-with-order}.
Note that if we omit the \textbf{QE-exp} step,
the sub-cases are equivalent to Cooper's algorithm,
so we adopt the idea of Oppen's analysis to analyze the complexity
\cite{Oppen}.

Apply Cooper's algorithm on the quantified PA formula 
$Q_m x_m Q_{m-1}$ 
$x_{m-1} ... Q_1 x_1. $
$F(x_1,...,x_m)$,
the algorithm repeats $m$ times to eliminate
$x_1,...,x_m$ one by one.
Let $F_k \Def Q_m x_m Q_{m-1} 
x_{m-1} ... Q_{k+1} x_{k+1}. F_k'(x_{k+1},...,x_m)$
be the formula produced after $k$th iterations of the algorithm. 
Let $a_k$ denote the number of atoms in $F_k$,
$c_k$ denote the number of distinct $\delta_i$ in atom $\delta_i | t$ (and its negation) plus the number of distinct coefficients of variables.
$s_k$ denote the largest value of integer constant.
Use $a_0,c_0,s_0$ to denote the initial value of $a_k,c_k,s_k$ respectively.
The following theorem holds.

\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {s_{k-1}}^{4c_{k-1}}$
    \item $a_k \le {a_{k-1}}^4 \cdot {s_{k-1}}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

For each $k$,
the space required to store $F_k$ 
is $a_k \cdot (m+1) \cdot s_k \cdot q$,
where $(m+1)$ is the maximum number of constants per atom and $q>1$ is some constant. Assume $m\le n, c\le n, a\le n, s\le n$, we obtain the deterministic space complexity is ${{2^2}^2}^{p n log_n}$, which is also the bound for deterministic time.

In our algorithm,
we use the same denotations and have the following theorem. 
Note that each iteration corresponds to \textbf{QE-exp} and \textbf{QE-linear}.
\begin{theorem}
for all $k: 1\le k\le m$, the following relations holds
\begin{itemize}
    \item $c_k \le {c_{k-1}}^4$
    \item $s_k \le {(n s_{k-1})}^{4c_{k-1}}$
    \item $a_k \le {{(l_2(ns_{k-1})\cdot a_{k-1})}}^4 \cdot {(n s_{k-1})}^{2c_{k-1}}$
\end{itemize}
and from the above relations we have 
\begin{itemize}
    \item $c_k \le {{c_0}^4}^k$
    \item $s_k \le {{{n}^{4c_0}}^4}^k \cdot {{{s_0}^{4c_0}}^4}^k$
    \item $a_k \le {{a_0}^4}^k \cdot 
     {{{n}^{4c_0}}^4}^k
     \cdot {{{s_0}^{4c_0}}^4}^k$
\end{itemize}
\end{theorem}

WAIT TO CHECK!

Again assume $m,c,a,s\le n$, which gives 
$$
\textit{space}\le q \cdot {n^4}^n \cdot {{n^{(4n)}}^4}^n \cdot
{{n^{(4n)}}^4}^n \cdot (n+1) \cdot 
\cdot {{n^{(4n)}}^4}^n \cdot {{n^{(4n)}}^4}^n
\le 2^{2^{2^{p n \textit{log}n}}}
$$

Then the upper bound for Cooper's algorithm still holds for the inner \textit{for} loop of \textbf{QE-with-order}.
When considering \textbf{Normalization} and permutations groups in \textbf{QE-with-order},
since the super exponential term dominates the number of permutations $n^{O(n)}$,
this upper bound still holds.

\subsection{Example}

We apply our algorithm to a simple example to 
illustrate its function.
Given $\Sigma = \{0,1\}$
Suppose $\mathcal{A}$ is a finite state automaton
that recognizes language $L = \{ w \mid \#w('1') \text{ is odd}\}$.
Consider the following string constraint
$$y\in L(\mathcal{A})\wedge 4\le  \parseInt(y) < 16\,.$$

We need an abstraction parameter $alpha$ to restrict interpretations
for $x$, for simplicity, set $\alpha\Def \langle 2,1 \rangle$.
The generic $\langle 2,1 \rangle$-flat language is the union of 
$\langle 2,1 \rangle$-flat languages $(11)^*,(10)^*,(01)^*,(00)^*,1^*,0^*$.
Here we divide the problem into 6 sub-cases by restricting $y$
to one of the $\langle 2,1 \rangle$-flat language.
Another approach is to directly restrict $y$ to 
the generic $\langle 2,1 \rangle$-flat language,
at the cost of more introduced variables.
Since the complexity is related to the number of variables,
we wish it to be as little as possible.

When $y$ is in language $1^*$, assume $y = 1^x$ where $x$
is a fresh integer variable.
Note that the Parikh image of $L(\mathcal{A})$ is $2|( x+1)$, 
so the constraint is converted to a PA formula $\phi$
$$\phi\Def \exists x. 2|(x+1) \wedge 5 \le 2^{x} < 17\,.$$

Observe that when $x=3$ the formula holds,
so we can easily deduce that $\phi$ is true.  
By applying our algorithm,
after \textbf{Normalization},
$\phi$ is transformed into the form

$$\exists x_0. 2|(x+1)\wedge  2^{x}\le 16 \wedge -2^{x}\le -5\,.$$

Since there is only one variable, 
\textbf{QE-with-order} can be omitted.
We send $2^x\le 16$ and $-2^x \le -5$ to \textbf{QE-exp} and obtain
$$2^x\le 16 \equiv x \le 4$$
$$-2^x\le -5 \equiv 3\le x\,.$$ 
Then we apply \textbf{QE-linear} on formula $\exists x_0. 2|(x+1) \wedge x\le 4 \wedge 3\le x$ and get
$$\bigvee_{j=0}^1(2| j+1 \wedge \textit{false} \wedge \textit{true})\vee \bigvee_{j=0}^1 (2|4-j+i\wedge 4-j\le 4 \wedge 3\le 4-j) \,.$$
which evaluates \textit{true} and we can deduce $j=1\implies x=3$, so $y$ is string $111$.

Other sub-cases are similar. However, when the abstraction parameter becomes larger,
say, $\alpha =\langle 2,2 \rangle$,
the complexity will grow rapidly,
because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.

\section{Conclusion and Future Works}

%test bib
\bibliographystyle{plain}
\bibliography{reference}

\end{document}