### 2021-05-21 FMCAD2021/7_conclusion.tex
conclusion

%!TEX root = paper.tex

In this paper, we proposed a complete flattening approach for string constraints with string-integer conversion and flat regular constraints, based on a quantifier elimination procedure by Point in 1986, for the extension of Presburger arithmetic with exponential functions. We gave a more accessible reformulation of Point's procedure and proposed various optimizations. Moreover, we achieved the first prototypical implementation of Point's procedure. We also did extensive experiments to evaluate the performance of the implementation. The experiment results show the efficacy of our implementation, compared with the state-of-the-art solvers. An interesting question left open in this paper is a more refined complexity analysis of the quantifier elimination procedure, considering the fact that quantifier elimination procedure for Presburger arithmetic, e.g. Cooper's algorithm, is triply exponential.